Supplementary Material for 'New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-16 Manuscript/Manuscript.Rmd
Finished Sup Material

Finally, we run extinction simulations and estimated its impact on the stability of the food web. For this, we calculated a stability index called Quasi-Sign Stability (QSS), which is the proportion of stable networks using randomized Jacobians and keeping the predator-prey sign structure fixed [@Allesina2008]. If this proportion is zero, then one should take into account the real part of the maximum eigenvalue of the Jacobian matrix, which is also a measure of stability [@Grilli2016].

@@

It's important to note that since the proportion of Jacobians that are locally stable or QSS was zero for the Weddell Sea food web, we considered the mean maximum eigenvalue as the stability index. When performing the extinction simulations, our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Figure 5). This is shown in Figure 5, where most of the points lie around the zero value of the 'Stability difference' y axis, meaning that the stability for the food web minus a given species is similar (not significant) to the stability for the whole network. However, there were few species that when become extinct altered significantly the stability of the food web. Most of such species generated positive significant differences in the stability of the food web, which mean that they increased the network's stability when deleted. There were two exceptions, Pagetopsis macropterus (demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod), that when deleted decreased the stability. Overall, 15 out of 490 species (3.06%) gave rise to significant changes in the food web's stability (Table 2).

>>>>

Finally, we run extinction simulations and estimated its impact on the stability of the food web. For this, we calculated a stability index called Quasi-Sign Stability (QSS), which is the proportion of stable networks using randomized Jacobians and keeping the predator-prey sign structure fixed [@Allesina2008]. If this proportion is zero, then one should take into account the real part of the maximum eigenvalue of the Jacobian matrix, which is also a measure of stability [@Grilli2016], indicating a more stable food web when it is negative.

@@

It's important to note that since the proportion of Jacobians that were locally stable or QSS was zero for the Weddell Sea food web, we considered the mean maximum eigenvalue as the stability index. When performing the extinction simulations, our results showed that the majority of the species had no significant impact on the stability of the food web when extinct (Figure 5). This is shown in Figure 5, where most of the points lie around the zero value of the 'Stability difference' y axis, meaning that the stability for the food web minus a given species is similar (not significant) to the stability for the whole network. However, there were few species that when become extinct altered significantly the stability of the food web. Most of such species generated positive significant differences in the stability of the food web, which mean that they increased the network's stability when deleted. There were two exceptions, Pagetopsis macropterus (demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod), that when deleted decreased the stability. Overall, 15 out of 490 species (3.06%) gave rise to significant changes in the food web's stability (Table 2).

### 2022-11-16 Manuscript/SupMaterial.Rmd
Finished Sup Material

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Table S1 shows the mentioned properties for every species of the Weddell Sea food web.

```{r echo=FALSE, results='asis'}
library(knitr)
library(dplyr)
load("net_&_spp_prop.rda")
knitr::kable(spp_all_prop %>% dplyr::select(TrophicSpecies, IS_mean, TotalDegree, TL, meanTrophicSimil) %>% mutate_if(is.numeric, round, digits=4))
```

@@

We performed extinction simulations, one at a time, for every species in the Weddell Sea food web. In order to assess the impact on the stability of the food web we statistically compared a stability index before and after performing the extinction. For this, we applied Quasi-Sign Stability $QSS$ that calculates the proportion of matrices that are locally stable. These matrices are created by sampling the values of the community matrix (the Jacobian) from a uniform distribution, preserving the sign structure: positive for predators and negative for prey. This stability index was originally proposed by @Allesina2008.
We used the R package multiweb to calculate $QSS$ and to test the $QSS$ difference before and after performing the extinction (Saravia, 2019). Two functions were specifically created for these analyses: ‘calc_QSS’ and ‘calc_QSS_extinction_dif’.
For the $QSS$ calculation we used a uniform distribution between 0 and maximum values given by the parameters negative, positive and self-damping, corresponding to the sign of interactions and self-limitation effect. Since we had estimated the interaction strength for each interaction of the Weddell Sea food web, the limits of the distribution were $negative*-x$,$positive*x$, $self-damping*x$, where $x$ is the value of the strength for the interaction in question.

Table S2 shows the results for the QSS differences before and after performing extinction simulations for every species of the Weddell Sea food web.

<!-- knitr::kable(websTbl %>% dplyr::select(Network,Size,Links,Connectance,PathLength,Clustering,SWness,SWnessCI,Q,zQ,Qlow,Qhigh,mTI,zTI,TIlow,TIhigh,Modularity,zMO,MOlow,MOhigh) %>% mutate_if(is.numeric, round, digits=4)) -->

>>>>

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

@@

Table 1 shows the mentioned properties for every species of the Weddell Sea food web.

```{r echo=FALSE, results='asis'}
library(knitr)
library(dplyr)
load("QSS_summary_sep22.rda")
knitr::kable(all_data %>% dplyr::select(TrophicSpecies, IS_mean, TotalDegree, TL, meanTrophicSimil) %>% arrange(., desc(IS_mean)), digits = c(0,7,0,2,3), col.names = c("Species","IS_mean", "Degree", "TL", "TS"), caption = "Weighted (interaction strength) and unweighted properties of the species of Weddell Sea food web. Ordered by decreasing mean interaction strength. IS_mean = mean interaction strength, TL = trophic level, TS = trophic similarity.")
```

@@

We performed extinction simulations, one at a time, for every species in the Weddell Sea food web. In order to assess the impact on the stability of the food web we statistically compared a stability index before and after performing the extinction. For this, we applied Quasi-Sign Stability $QSS$ that calculates the proportion of matrices that are locally stable. These matrices are created by sampling the values of the community matrix (the Jacobian) from a uniform distribution, preserving the sign structure: positive for predators and negative for prey. This stability index was originally proposed by @Allesina2008.
We used the R package multiweb to calculate $QSS$ and to test the $QSS$ difference before and after performing the extinction (Saravia, 2019). Two functions were specifically created for these analyses: ‘calc_QSS’ and ‘calc_QSS_extinction_dif’.
For the $QSS$ calculation we used a uniform distribution between 0 and maximum values given by the parameters negative, positive and self-damping, corresponding to the sign of interactions and self-limitation effect. Since we had estimated the interaction strength for each interaction of the Weddell Sea food web, the limits of the distribution were $negative*-x$,$positive*x$, $self-damping*x$, where $x$ is the value of the strength for the interaction in question.
We performed 1000 extinction simulations for every species. Our results showed that the proportion of Jacobians that were locally stable was zero. Thus, we considered the mean maximum eigenvalue as the stability index, hereafter $QSS$. For testing the $QSS$ difference before and after the extinction we performed an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987].

Table 2 summarizes the $QSS$ results for every species extinction of the Weddell Sea food web.

```{r echo=FALSE, results='asis'}
library(knitr)
library(dplyr)
load("QSS_summary_sep22.rda")
knitr::kable(all_data %>% dplyr::select(TrophicSpecies, QSS_all, QSS_ext, difQSS, Ad_pvalue) %>% arrange(., Ad_pvalue), digits = c(0,7,7,7,41), col.names = c("Species", "QSS_all", "QSS_ext", "difQSS", "AD_pvalue"), caption = "Summary of Quasi-Sign Stability (QSS) results before and after performing extinction simulations in the Weddell Sea food web. Ordered by increasing p-values of the Anderson-Darling test. QSS_all = Quasi-Sign Stability before the extinction (whole food web), QSS_ext = Quasi-Sign Stability after the extinction, difQSS = QSS difference between 'QSS_all' and 'QSS_ext', AD_pvalue = p-value for the Anderson-Darling test.")
```
